Structural-temporal neighbourhoods of posts to characterize conversations in online forums

### 2016-04-29 doc/IEEEtran/paper_IEEEformat.tex
paper almost final

\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{algorithm}
\newtheorem{definition}{Definition}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

@@

Rather than looking for a fixed time radius, we may instead decide it by looking at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation. If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
\end{definition}
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.

% Explain algorithm
Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the given definition. Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree. In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches. Once we have the breakpoints we can proceed with the algorithm. First, we extract the structural-neighbourhood. The time-based neighbourhood will be a subset of the later. Then, we look for horizontal and vertical breakpoints, which mark the frontiers of the time-based neighbourhood. There are four possible cases:

\begin{itemize}
\item A \textit{vertical breakpoint in the ancestors}:  If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs. Thus, we remove the ancestors of the breakpoint, but not the breakpoint.
\item A  \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself. 
\item An \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs. Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic. In both cases, we remove the older siblings of the breakpoint but not the breakpoint.
\item An \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
\end{itemize}
Any time a vertex is remove we also remove its descendants. The final neighbourhood will be connected induced subgraph from the structural neighbourhood. If there are no changepoints in the structural neighbourhood, the time-base neighbourhood will be exactly the structural neighbourhood.

@@

\section{Neighbourhood colouring and pruning}\label{sec:colouring_pruning}
\subsection{Colouring}
Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation. We can easily reduce this ambiguity by assigning colors, or labels, to vertices, identifying some relevant property of the post.

@@

\begin{itemize}
\item \textit{Red}: ego post.
\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (the same author participating more times in the discussion \cite{Backstrom2013})
\item \textit{Yellow}: parent of ego post and other posts written by the same author. It allows to identify, for instance, debates between the ego author (red and orange) and the other author (yellow). We have observed this phenomena in our data, and it is often the cause of long chains.
\item \textit{White}: root post. Differentiating the root post from the rest has been proven by previous research on online discussions. For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}). Also, in terms of preferential attachment, root posts use to get more replies than the non-roots \cite{Gomez2010, Gomez2012}.
\item \textit{Black}: none of the above.
\end{itemize}
Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{black, orange, red, yellow, white}$.

We do not claim this choice of labels to be of universal. Indeed, other labellings might also give interesting results since they would look at conversations from new points of view (e.g.: a colour for leaf vertices, colours according to the length of the post, or colours to represent the sentiment of the post).

\begin{figure*}
\centering
\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
\includegraphics[width=0.19\textwidth]{neighbourhood_time_456}
\includegraphics[width=0.18\textwidth]{neighbourhood_time_501}
\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
\caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
\label{fig:colors}
\end{figure*}

\subsection{Pruning}
At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.    
Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.

\section{Application to Reddit forums}
Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014. Nowadays, its members use it mainly to share and discuss about political news. The Game of Thrones is a casual discussion forum about the Game of Thrones TV series.
The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between the April 25th and October 20th 2014. The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.

\subsection{Neighbourhood extraction}
As a previous step for all further analysis we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums. We label every structure with a unique integer number, where a structure is a coloured graph. Two neighbours have the same structure, and therefore they have the same integer number, if their coloured graphs are isomorphic.

Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data. It appears then a much more practical approach to enumerate only those trees that are seen in the data. In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings) we re-label the structures after the extraction for every post is finished. We will call this final labelling a \textit{dictionary} of neighbourhoods.

Given a type of neighbourhood (struture-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows. A counter for every neighbourhood structure is created the first time it us detected. For every post, we extract its neighbourhood, colour and prune it (Sections~\ref{sec:structural} and \ref{sec:temporal}). Then we check whether it is isomorphic to some of the neighbourhoods that we have already seen. If this is the case, then we increment the counter for the neighbourhood previously seen. Otherwise we create a new entry for the current neighbourhood.

Once the six extractions are finished we merge there dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth). Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.

% Computational cost
In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood in already in the dictionary before concluding that it needs a new entry. In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts. In practice, however, the bigger the number of posts processed, the less likely that we go through all the dictionary without finding and isomorphism. Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find and isomorphism among the first neighbours of the dictionary.

\subsection{Comparing neighbourhood methods}
In this section we compare different aspects of our three classes of neighbourhood, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, (c) the neighbourhood census obtained for each class in a same forum.

\subsubsection{Size and frequency distribution}
We detected, in the the whole dataset, 3791 different structure-based neighbourhoods, 174 different order-based and 1488 different time-based. which means that the structure-based captures a much wider set of structures. Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods (41\%) and the structure-based neighbourhoods (60\%) occur only once, while the number of neighbourhoods that appear more than 100 times (among 150,000 posts) is similar for the three  types of neighbourhood. To cover 95\% of the posts we need 623 structure-based neighbourhoods, 40 order-based and 155 time-based.

% Why order-based is not good enough
The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id 1) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood. This is the main handicap of the order-based neighbourhood since, as we will see later, it avoids finding richer structures.

@@

\subsubsection{Discrepancies}
The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, in time-based, might be systematically mapped to the same neighbourhoods in order-based. Or similarly, that some structure-base neighbourhoods are systematically mapped into the same time-based neighbourhoods. Figure~\ref{fig:confusion} gives a general overview of these mappings. The left figure shows, for every post, its time-based neighbourhood and its order-based neighbourhood, while the right figure shows the time-based and the structure-based. The points in the diagonal correspond to posts that have been assigned the same neighbourhood by the two compared methods.

The diagonal of the time-based / order-based plot contains 36\% of posts. The diagonal of the structure-based / time-based plot contains 40\% of the posts. For the order-based / struct-based (not shown) there are 41\% of posts in the diagonal. In general, the structures with bigger agreement are among the most frequent. This is comprehensible since the most frequent tend to be also the most simple structures with less room for discrepancy. There is a 30.75\% of posts that are assigned the same neighbourhood structure by the three methods.

Rather than looking at where the methods agree, it is even more interesting to see where they disagree, since this will reveal whether one of the methods is incapable of detecting some interesting conversational structure. In Figure~\ref{fig:confusion} some horizontals lines can be spotted specially in the lower part of the plots. These correspond to posts with different time-based neighbourhoods that are assigned the same order-based neighbourhood (left) and different time-based neighbourhoods that are assigned the same order-based neighbourhood (right). Looking for the order-based neighbourhoods that have a correspondence with a larger set of time-based neighbourhoods, we detect that these correspond to neighbourhoods 1 (77 different time-based) and 50 (57 different time-based), 17 and 18. Looking for the time-based neighbourhoods that have a correspondence with a larger set of structure-base neighbourhoods, we detect that these are neighbourhoods 10, 4, 8, 19, 14.

% Wrap-up
In retrospective, these results are not surprising. What we see is simply that structures that are often captured by one method collapse into another single structure when other method with a smaller space of structures is used. Yet, the three methods mostly agree when they find a frequent neighbourhood. But this might be due to the fact that simpler motifs tend to be in shorter threads, where there is no room for two neighbourhood methods to detect a different structure.

@@

\subsubsection{Neighbourhood census}
Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph. In the context of online discussion forums, triad census has been sometimes used in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}. However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply). On the contrary, our neighbourhoods are potentially richer structures that respect the real structure of conversations.

In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type. Because some of the most frequent neighbourhoods in one method are also among the most frequent in the other methods, the total set contains 30 different neighbourhoods.

Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones. These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25). The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ... (27), suggesting that this method fails to capture some relevant structures.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{structure_based_census}
\includegraphics[width=0.5\textwidth]{order_based_census}
\includegraphics[width=0.5\textwidth]{time_based_census}
\caption{Comparison of neighbourhood census obtained by each type of neighbourhood in two different forums. Order-based is more similar across forums since its most frequent neighbourhood dominates the census.}
\label{fig:census_compare}
\end{figure}

We conclude that, even if at the cost of a sparser set of structures, the time-based neighbourhood does a better job in capturing structures that have both a real meaning and a level of frequency that makes them relevant in the analysis.

@@

Our dataset contains a set of 100 \textit{active users} who wrote more than 100 posts. For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$. We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity). To make the feature vector of a user independent on the number of posts, we transform the counts into percentages. And since some features have much higher percentages than others for most users, we scale and normalize the matrix so that every feature has mean 0 and variance 1. To avoid non-significant scores, we remove also the feature column corresponding to neighbourhoods that have a frequency less than 50 among the active users.

@@

% TODO: paragraph discussing a visual/manual analysis of some users

@@

% Maybe mention that we can take for instance only the posts before the ego and use it to understand when when can recommend a post based on the structure.

\section{Conclusions}
We presented a method to characterise conversations of users in online threads. Due to the tree nature of online threads, traditional patters such as triads are not able to capture much of relevant dynamics of a conversation. Our defined order-based and temporal-based neighbourhoods 
are able to capture a very rich variety of structures. We used this neighbourhoods to characterise users in terms of the structure of the conversations they participate in and showed that, indeed, there are different types of structural conversationalists

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\newpage
\section*{Appendix  (WILL ME REMOVED)}

\begin{figure*}
\centering
%\includegraphics[width=0.5\textwidth]{time_neighbourhood_census_compare_forums}%
%\includegraphics[width=0.5\textwidth]{order_neighbourhood_census_compare_forums}
\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_gameofthrones}%
\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_podemos}
\caption{Comparison of census for two different forums. POUR BERTRAND. WE CANT SEE ANYTHING HERE}
\label{fig:census_compare}
\end{figure*}

\begin{figure*}
	\centering
%	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
%	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$ (Podemos)}
	\label{fig:census_orderbased}
\end{figure*}
\begin{figure*}
	\centering
%	\includegraphics[width=0.8\textwidth]{census_timebased_1}
%	\includegraphics[width=0.8\textwidth]{census_timebased_2}
	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
	\caption{Most frequent time-based neighbourhoods (Podemos)}
	\label{fig:neighbourhoods_time}
\end{figure*}

\begin{figure*}
\centering
%\includegraphics[width=1\textwidth]{confusion_time_order_motifs_gameofthrones_1}
\caption{Most popular time-based neighbourhoods (up) and how they are mostly detected in order-based neighbourhood (down).}
\label{fig:census_size_vs_freq}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
	\label{fig:PCA}
\end{figure*}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{multirow}
\newtheorem{definition}{Definition}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

@@

Rather than looking for a fixed time radius, we may instead decide it by looking at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation. If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}.}. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
\end{definition}
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.

% Explain algorithm
Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the above definition. Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree. In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches. Once we have the breakpoints, we can proceed with the algorithm. First, we extract the structural-neighbourhood. The time-based neighbourhood will be a subset of the later. Then, we look for horizontal and vertical breakpoints, which mark the frontiers of the time-based neighbourhood. There are four possible cases:

\begin{itemize}
\item A \textit{vertical breakpoint in the ancestors}: If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs. Thus, we remove the ancestors of the breakpoint, but not the breakpoint itself.
\item A \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself. 
\item A \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs. Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic. In both cases, we remove the older siblings of the breakpoint but not the breakpoint itself.
\item A \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
\end{itemize}
Any time a vertex is removed we also remove its descendants. The final neighbourhood will be connected induced subgraph from the structural neighbourhood. If there are no changepoints in the structural neighbourhood, the time-based neighbourhood will be exactly the structural neighbourhood.

@@

\section{Neighbourhood colouring and pruning}\label{sec:colouring_pruning}
\subsection{Colouring}
Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation. We can easily reduce this ambiguity by assigning colours to vertices, which allows us to identify some relevant property of the post.

@@

\begin{itemize}
\item \textit{Red}: ego post.
\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (when the same author participates several times in the discussion \cite{Backstrom2013})
\item \textit{Yellow}: parent of ego post and other posts written by the same author. It allows to identify, for instance, debates between the ego author (red and orange) and the other author (yellow). We have observed this phenomena in our data, and it is often the cause of long chains.
\item \textit{White}: root post. Differentiating the root post from the rest has been proven by previous research on online discussions. For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}). Also, in terms of preferential attachment, root posts usually get more replies than the non-roots (\cite{Gomez2010, Gomez2012}).
\item \textit{Black}: none of the above.
\end{itemize}
Since some posts might correspond to several colours, we perform a sequential assignment of colours given by $\texttt{black, orange, red, yellow, white}$.

We do not claim this choice of labels to be of universal. Indeed, other labellings might also give interesting results since they would look at conversations from new points of view (e.g.: a colour for leaf vertices, colours according to the post length, or colours to represent the sentiment of the post).

\subsection{Pruning}
At this point, we can still find structures the only difference of which is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same colour and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.

\section{Application to Reddit forums}
Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid on October 18th and 19th, 2014. Nowadays, its members use it mainly to share and discuss about political news. The Game of Thrones forum is a casual discussion forum about the Game of Thrones TV series.
The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between April 25th and October 20th, 2014. The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.

\subsection{Neighbourhood extraction}
As a previous step for all further analysis, we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums. We label every structure with a unique integer number, where a structure is a coloured graph. Two neighbours have the same structure, and therefore the same integer number, if their coloured graphs are isomorphic.

Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of these trees would not be seen in the data. It appears then a much more practical approach to enumerate only those trees that are seen in the data. In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished. We will call this final labelling a \textit{dictionary} of neighbourhoods.

Given a type of neighbourhood (structure-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows. A new label is assigned to every neighbourhood structure the first time it is detected. For every post, we extract its neighbourhood, colour and prune it (Sections~\ref{sec:structural} and \ref{sec:temporal}). Then we check whether it is isomorphic to some of the $n$ neighbourhoods that we have already seen. If this is the case, then the current neighbourhood is given the same label than the neighbourhood that we already saw. Otherwise we create a new label for the current neighbourhood.

Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth). Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.

% Computational cost
In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood that already appears in the dictionary before concluding that it needs a new entry. In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts. In practice, however, the bigger the number of posts processed, the less likely we go through all the dictionary without finding an isomorphism. Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find an isomorphism among the first neighbours of the dictionary.

\subsection{Comparing neighbourhood methods}
In this section we compare different aspects of our three classes of neighbourhoods, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, and (c) the neighbourhood census obtained for each class in the same forum.

\subsubsection{Size and frequency distribution}
We detected in the the whole dataset 3,791 different structure-based neighbourhoods, 174 different order-based and 1,488 different time-based, which means that the structure-based captures a much wider set of structures. Of course, this is not necessarily an advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods (41\%) and the structure-based neighbourhoods (60\%) occur only once, while the number of neighbourhoods that appear more than 100 times (among 150,000 posts) is similar for the three types of neighbourhood. To cover 95\% of the posts we need 623 structure-based neighbourhoods, 40 order-based and 155 time-based.

% Why order-based is not good enough
The main reason why the frequency of the order-based neighbourhoods decays faster is that the most frequent neighbourhood (id 1) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood. This is the main handicap of the order-based neighbourhood since, as we will see later, it avoids finding richer structures.

@@

\begin{table}
 \begin{center}
   \tabcolsep = 0.8\tabcolsep
   \begin{tabular}{|cc|cc|cc|}
      \hline
      \multicolumn{2}{|c|}{struct-based} & \multicolumn{2}{|c|}{order-based} &  \multicolumn{2}{|c|}{time-based} \\
      %\hline
     	 POD & GOT 						& POD & GOT 		& POD & GOT\\
   \hline \hline
     1 (12.5)& 3(12.5)   &  1 (12.5) & 1 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  2 (12.5) & 2 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\  
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\          
   \hline
   \end{tabular}
\label{tab:table}
 \end{center}
 \caption{Frequence}
\end{table}

\subsubsection{Discrepancies}
The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, time-based, might be systematically mapped to the same neighbourhoods in order-based. Or similarly, that some structure-based neighbourhoods are systematically mapped into the same time-based neighbourhoods. Figure~\ref{fig:confusion} gives a general overview of these mappings. The left figure shows, for every post, its time-based neighbourhood and its order-based neighbourhood, while the right figure shows the time-based and the structure-based. The points in the diagonal correspond to posts that have been assigned the same neighbourhood by the two compared methods.

The diagonal of the time-based / order-based plot contains 36\% of the posts. The diagonal of the structure-based / time-based plot contains 40\% of the posts. For the order-based / struct-based (not shown) there are 41\% of the posts in the diagonal. In general, the structures with a bigger agreement are among the most frequent. This is comprehensible since the most frequent tend to be also the most simple structures with less room for discrepancy. There is a 30.75\% of the posts that are assigned the same neighbourhood structure by the three methods.

Rather than looking at where the methods agree, it is even more interesting to see where they disagree, since this will reveal whether one of the methods is incapable of detecting some interesting conversational structure. In Figure~\ref{fig:confusion} some horizontals lines can be spotted specially in the lower part of the plots. These correspond to posts with different time-based neighbourhoods that are assigned the same order-based neighbourhood (left), and different time-based neighbourhoods that are assigned the same order-based neighbourhood (right). Looking for the order-based neighbourhoods that have a correspondence with a larger set of time-based neighbourhoods, we detect that these correspond to neighbourhoods 1 (77 different time-based) and 50 (57 different time-based), 17 and 18. Looking for the time-based neighbourhoods that have a correspondence with a larger set of structure-based neighbourhoods, we detect that these are neighbourhoods 10, 4, 8, 19, 14.

% Wrap-up
In retrospective, these results are not surprising. What we see is simply that structures that are often captured by one method collapse into another single structure when another method with a smaller space of structures is used. Yet, the three methods mostly agree when they find a frequent neighbourhood. But this might be due to the fact that simpler motifs tend to be in shorter threads, where there is no room for two neighbourhood methods to detect a different structure.

@@

It is clear from the last sections that the order-based does not do a good job on capturing all the richness of the conversational structures. There are two other reasons to choose the time-neighbourhood over the structure based. First, the time-based has less spurious structures, covering 95\% of posts with 155 neighbourhoods, against the 623 neighbours necessary to cover that amount with the structure-based. Second, the time-based neighbourhood has more grounded criteria to decide where the neighbourhood ends, while the structure-based has no criteria at all.

@@

Our dataset contains a set of 100 \textit{active users} who wrote more than 100 posts. For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the dictionary, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$. We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity). To make the feature vector of a user independent on the number of posts, we transform the counts into percentages. And since some features have much higher percentages than others for most users, we scale and normalize the matrix so that every feature has mean 0 and variance 1. To avoid non-significant scores, we remove also the feature column corresponding to neighbourhoods that have a frequency less than 50 among the active users.

@@

@@

\section{Conclusions}
The goal of is paper was to shed some light on what would be a proper method to characterise the discussion in which a post is embedded. For that, we proposed three different kinds of neighbourhood applicable to tree structures by which conversation threads are represented. We named this three types of neighbourhood structure-based, order-based and time-based. While the first one is a mere extension of the classical concept of neighbourhood in a graph, the aim of the other two is to explicitly consider time when deciding which are the neighbours of a given post.

We showed the limitations of the order-based to capture the structure of conversations. We think this finding is relevant because, in some contexts, order is more relevant than real time, and assuming that posts (or whatever vertices represent) arrive in homogeneous intervals gives excellent results \cite{Gomez2012}.

We also showed that the main contribution of the time-based neighbourhood with respect to the most basic structure-based is that, thanks to the use of changepoints to set limits of the neighbourhoods, the space of neighbourhoods is considerably reduced and with less occasional neighbourhoods that occur only once. Besides, it prevents neighbourhoods from being too big, even though some post-pruning is still necessary.

We analysed the time .......

 the  it is common in the literature to use the order rather than the real time. The

We introduced three different  a method to characterise conversations of users in online threads. Due to the tree nature of online threads, traditional patters such as triads are not able to capture much of relevant dynamics of a conversation. Our defined order-based and temporal-based neighbourhoods 
are able to capture a very rich variety of structures. We used this neighbourhoods to characterise users in terms of the structure of the conversations they participate in and showed that, indeed, there are different types of structural conversationalists

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography}
